FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-05-08 2018EA000388_Answers_1Reviewer.txt

2018EA000388 Answers to the first Reviewer

l. 43
Rephrased: see the new version of the manuscript

l. 57
We agree with the reviewer: the paragraph has been deleted

l. 68
Rephrased: see the new version of the manuscript

l. 68-69
IPDA informally discussed about proposing recommended data formats for archive.
As an IPDA official position is missing we cannot argue about FITS in that context.

l. 83-84

l. 102
GDAL lack of metadata description for FITS format: added in the manuscript

Table 1: note b

Table 2 : header
Mercury must be there... sorry for the oversight.
TNO and KBO are Asteroids: this is now clarified in the note.

### 2018-05-08 ess_marmo.tex
Working on ESS first internal review.

\section{Introduction}
\label{sec:intro}
Planetary science encompasses a broad number of research fields and
brings together several research communities (geologists, astronomers,
physicists, geochemists, etc.). Planetary missions produce an impressively growing 
amount of diverse data requiring an evolution from a mostly manual-based visual 
analysis to a more automated and quantitative analysis.
Interoperability, openness of data formats, and shared processing techniques are 
a necessity to efficiently extract scientific information from the data and 
to guarantee the reproducibility of the scientific results.

@@

This paper is organized as follows:
\begin{enumerate}
\item{a brief description of FITS format;}
\item{a proposition for planetary surface metadata description in FITS;}
\item{dictionaries between FITS, PDS metadata description and EPN-TAP2.0
parameters;}
\item{some examples of planetary dataset in FITS;}
\item{current developments for making FITS more usable to handle data
of planetary surfaces.}
\end{enumerate}

@@

\subsection{FITS Format}
FITS is an open digital standard, created in the late 1970's for data acquisition,
transfer and archiving of telescope data by astronomical observatories.
It had been used for data exchange and archiving from several orbital telescopes.
The International Astronomical Union approved FITS as the standard format
for astronomical data\footnote{https://fits.gsfc.nasa.gov/iaufwg/history/IAU\_1982\_resolution\_c1.html}. 
Therefore, FITS is one of the standard formats implemented in the Virtual
Observatory (VO).
FITS data storage is compatible with the Planetary Data System (PDS)
archiving specifications so that FITS files can be embedded in PDS datasets.
Finally, it is supported by a large number of open libraries
and software tools.

@@

\subsection{Planetary Surface Proposed Convention}
\label{ssec:geofits}
To facilitate surface investigations, data taken of a planetary surface should be
radiometrically calibrated and orthorectified (when topographic data is available)
into a well-known map projection as required by Geographical Information Systems (GIS)
applications.
In general, GIS applications excel in planetary data set integration and 
interoperability even though some have historically been anchored to Earth's spatial
description.
Often, robust interoperability in GIS applications is achieved by using the Geospatial
Data Abstraction Library\footnote{http://www.gdal.org} (GDAL), released by the Open
Source Geospatial Foundation.
GDAL is essentially a format translation library written in C++ for geospatial raster
and vector data and, fortunately, it also offers basic support for the FITS format.

@@

\begin{table}
\caption{Codes to be used in CTYPE FITS keyword in order to identify the planetary body.
Codes for planets are assumed in \citet{calagreisenI}.
We propose to add codes for classes of targets,
more or less corresponding to PDS4 Target type attribute and to EPN-TAP
target\_class parameter.}
\centering
\begin{tabular}{l l}
\hline
Body & CTYPE$^{a}$ \\ 
\hline
Moon & SE \\
Venus & VE \\
Mars & MA \\
Jupiter & JU \\
Saturn & SA \\
Uranus & UR \\
Neptune & NE \\
Satellites (other than the Moon) & ST \\
Asteroids & AS \\
Dwarf Planets & DW \\
Comets & CO \\
\hline
\multicolumn{2}{l}{$^{a}$From \citet{calagreisenI}}
\end{tabular}
\label{table:ctype}
\end{table}

>>>>

\section{Introduction}
\label{sec:intro}
Planetary science encompasses a broad number of research fields and
brings together several research communities (geologists, astronomers,
physicists, geochemists, etc.). Planetary missions produce an impressively growing 
amount of diverse data requiring an evolution from a mostly manual-based visual 
analysis to a more automated and quantitative analysis.
Interoperability, openness of data formats, and shared processing techniques are 
a necessity to allow efficient scientific analysis of the data and 
to guarantee the reproducibility of the results.

@@

@@

\subsection{FITS Format}
FITS is an open digital standard, created in the late 1970's for data acquisition,
transfer and archiving of telescope data by astronomical observatories, where
it has been used during the last thirty years.
It had been adopted for data exchange and archiving from several orbital telescopes
and spatial missions.
The International Astronomical Union approved FITS as the standard format
for astronomical data\footnote{https://fits.gsfc.nasa.gov/iaufwg/history/IAU\_1982\_resolution\_c1.html}. 
Therefore, FITS is one of the standard formats implemented in the Virtual
Observatory (VO).
FITS data storage is compatible with the Planetary Data System (PDS)
archiving specifications so that FITS files can be embedded in PDS datasets.
Finally, it is supported by a large number of open libraries
and software tools.

@@

\subsection{Planetary Surface Proposed Convention}
\label{ssec:geofits}
To facilitate surface investigations, data taken of a planetary surface should be
radiometrically calibrated and orthorectified (when topographic data is available)
into a well-known map projection as required by Geographical Information Systems (GIS)
applications.
In general, GIS applications excel in planetary data set integration and 
interoperability even though some have historically been anchored to Earth's spatial
description.
Often, robust interoperability in GIS applications is achieved by using the Geospatial
Data Abstraction Library\footnote{http://www.gdal.org} (GDAL), released by the Open
Source Geospatial Foundation.
GDAL is essentially a format translation library written in C++ for geospatial raster
and vector data and, fortunately, it also offers basic support for the FITS format,
however missing a FITS metadata dictionary.

@@

\begin{table}
\caption{Codes to be used in CTYPE FITS keyword in order to identify the planetary body.
Codes for planets are assumed in \citet{calagreisenI}.
We propose to add codes for classes of targets,
more or less corresponding to PDS4 Target type attribute and to EPN-TAP
target\_class parameter.}
\centering
\begin{tabular}{l l}
\hline
Body & CTYPE$^{a}$ \\ 
\hline
Moon & SE \\
Mercury & ME \\
Venus & VE \\
Mars & MA \\
Jupiter & JU \\
Saturn & SA \\
Uranus & UR \\
Neptune & NE \\
Satellites (other than the Moon) & ST \\
Asteroids$^{b}$ & AS \\
Dwarf Planets & DW \\
Comets & CO \\
\hline
\multicolumn{2}{l}{$^{a}$From \citet{calagreisenI}} \\
\multicolumn{2}{l}{$^{b}$Transneptunian (TNO) and Kuiper Belt (KBO) objects must be placed there}
\end{tabular}
\label{table:ctype}
\end{table}
